Parallel for Oregon Zip and WRF_Grid Intersection Proportion

### 2017-03-28 r_scripts/parallel_presentation.Rmd
Update parallel_presentation.Rmd

```{r library setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(rgdal) # package for shape files
library(sp)
library(rgeos) # rgeos package contains the intersect and area commands I need
library(dplyr)
library(data.table)
library(readxl)
# parallel computing libraries
# library(foreach) 
library(doParallel)
library(parallel)
library(readr)
```

@@

## Shapefile preparation: 2013 Oregon Zip code Shapefile and WRF Grid
The shapefile of Oregon can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.
<br/>The whole U.S shapefile can be found from United States Census website:
<br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas
<br/>The Oregon zip code can also be found from: 
<br/>https://www.unitedstateszipcodes.org/or/

The plot below shows a Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.
```{r overlay}
plot(or_zip_map)
plot(smoke_grid)
plot(or_zip_map, add=T)
```

The Oregon shapefile contains some lack or extra, which may because the zip code changes (but U.S. shapefile is for 2013).

## Steps of calculating the proportion of a specific zip code polygon and WRF Grid.
1. Choose zip code = 94705. WRF Grid = 261 and 262. (6 WRF Grid in total. Only 
two are chosen for test)
```{r test, warning= FALSE}

test_zip <- c(97405)
test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]

@@

```{r include = FALSE}
test_grid <- over(smoke_grid, test_zip_map)
summary(test_grid) # 6 grids over the zip code
test_grid2 <- test_grid %>% filter(!is.na(ZCTA5CE10))
test_grid2
```

2. Show the actual location in the map.
```{r}
# test zip map over zip grid
plot(smoke_grid)
plot(test_zip_map, add = T)
zip_over_grid <- over(test_zip_map, smoke_grid)
# summary(zip_over_grid) # this way retains the values
```

@@

## Steps (Continued)
3. Show the intersection of chosen WRF Grid and zip code.
```{r show, warning=FALSE}
# Subset to one zip code and two different WRF grids ---------------------------
# plot zip and 2 grids
plot(test_zip_map)
#invisible(text(getSpPPolygonsLabptSlots(test_zip_map), 
#               labels=as.character(test_zip_map$ZCTA5CE10)))
plot(wrf_grid_261, add = T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
               labels=as.character(wrf_grid_261$WRFGRID_ID)))
plot(wrf_grid_262, add=T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_262), 
               labels=as.character(wrf_grid_262$WRFGRID_ID)))
```

4. The proportion of two WRF Grid.
```{r calculation, warning=FALSE}
# Start with the intersection with wrf grid 261 and zipcode
# first I need to convert the spatial polygon to just polygon
poly_261 <- SpatialPolygons(wrf_grid_261@polygons)
shape_zip <- SpatialPolygons(test_zip_map@polygons)

@@

<ul class="incremental"> 
  <li> One choice: <br/>  simple calculation; copy and paste</li> 
  <li> Another choice: for loop
  <br/> When analyzing big data set (like millions of observations), it is 
  unrealistic to copy so many rows of similar or same code.</li><br/>
  <li> And... <br/> Why using apply function? </li>
  <li> Code Readability <br/>  Understand the purporse more directly.</li> 
  <li> Working Efficiency <br/>  When analyzing big data set (like millions of 
  observations), it is unrealistic to copy so many rows of similar or same code.</li> 
</ul>

## Apply Family

- The apply() Function
<br/> Apply functions over array margins.<br/>

- The lapply() Function
<br/> Apply a function over a list or vector.<br/> 
Apply a given function to every element of a list and obtain a list as result.<br/>

- The sapply() Function
<br/> Apply a function over a list or vector.<br/>
Works like lapply(), but it tries to simplify the output to the most elementary data structure that is possible.

- The tapply() Function
<br/> Apply a function over a ragged array.

- The mapply() Function
<br/> Apply a function to multiple list or vector arguments.

## Parallel on server
One core vs Multiple cores

```{r}

for(i in 1:length(or_zip_name)){
  # output value of zipcode
  zipcode <- as.character(or_zip_name[i]) 
  # limit shapefile to particular zipcode
  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
  # convert to polygon
  zip_poly <-SpatialPolygons(zip_shape@polygons)

@@

>>>>

```{r librarysetup, include=FALSE}
library(rgdal) # package for shape files
library(sp)
library(rgeos) # rgeos package contains the intersect and area commands I need
library(dplyr)
library(data.table)
library(readxl)
# parallel computing libraries
# library(foreach) 
library(doParallel)
library(parallel)
library(readr)
```

@@

## Shapefile preparation: 2013 Oregon Zip code Shapefile and WRF Grid

Some libraries for spatial data analysis and parallel computing:
```{r, eval = FALSE}
library(rgdal)     # package for shape files
library(sp)        # spatial polygon
library(rgeos)     # rgeos package contains the intersect and area commands
library(parallel)  # parallel computing libraries
```

The **shapefile of Oregon** can be subset from the whole U.S. shapefile by filtering the zip code in Oregon State.

<br/>The **whole U.S shapefile** can be found from United States Census website:
<br/>https://www.census.gov/cgi-bin/geo/shapefiles/index.php?year=2013&layergroup=ZIP+Code+Tabulation+Areas

<br/>The **Oregon zip code** can also be found from: 
<br/>https://www.unitedstateszipcodes.org/or/

The plot below shows an Oregon State shapefile and an overlay of projected WRF Grid and Oregon shapefile.
```{r overlay, echo=FALSE}
plot(or_zip_map)
plot(smoke_grid)
plot(or_zip_map, add=T)
```

## Steps of calculating the proportion of intersection of a specific zip code polygon and WRF Grid.
1. Choose a zip code (94705 here) for testing. WRF Grid = 261 and 262. (6 WRF 
Grids in total. Only two are chosen for convenience)
```{r test, warning= FALSE, echo=FALSE}
test_zip <- c(97405)
test_zip_map <- or_zip_map[or_zip_map$ZCTA5CE10 %in% test_zip,]

@@

2. The actual location in the map.
```{r, echo=FALSE}
# test zip map over zip grid
plot(smoke_grid)
plot(test_zip_map, add = T)
zip_over_grid <- over(test_zip_map, smoke_grid)
# summary(zip_over_grid) # this way retains the values
```

@@

## Steps (Continued)
3. Show the intersection of chosen WRF Grid and zip code.
```{r show, warning=FALSE, echo=FALSE}
# Subset to one zip code and two different WRF grids ---------------------------
# plot zip and 2 grids
plot(test_zip_map)
#invisible(text(getSpPPolygonsLabptSlots(test_zip_map), 
#               labels=as.character(test_zip_map$ZCTA5CE10)))
plot(wrf_grid_261, add = T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_261), 
               labels=as.character(wrf_grid_261$WRFGRID_ID)))
plot(wrf_grid_262, add=T)
invisible(text(getSpPPolygonsLabptSlots(wrf_grid_262), 
               labels=as.character(wrf_grid_262$WRFGRID_ID)))
```

4. The proportion of two WRF Grid.
```{r calculation, warning=FALSE, echo=FALSE}
# Start with the intersection with wrf grid 261 and zipcode
# first I need to convert the spatial polygon to just polygon
poly_261 <- SpatialPolygons(wrf_grid_261@polygons)
shape_zip <- SpatialPolygons(test_zip_map@polygons)

@@

<ul class="incremental"> 
  <li> One choice: <br/>  simple calculation; copy and paste codes.</li> 
  <li> Another choice: for loop
  <br/> When analyzing big data set (for this data set, there are 417 zip codes 
  and 1610 WRF Grids in total), it is unrealistic to copy so many rows 
  (417*1610 = 671,370 times) of similar or same code.</li><br/>
  <li> And... <br/> Why using apply function? </li>
  <li> **Working Efficiency** <br/> **Vectorization.** Faster than for loop. It 
  works on margins of an array or matrix, not like for loop.</li> 
  <li> **Code Readability** <br/>  Understand the purporse more directly.</li> 
</ul>

## Apply Family
>- Returns a vector or array or list of values obtained by applying a function to 
margins of an array or matrix.

>- The apply() Function
<br/> Apply functions over array margins.<br/>

>- The lapply() Function
<br/> Apply a function over a list or vector.<br/> 
Apply a given function to every element of a list and obtain a list as result.<br/>

>- The sapply() Function
<br/> Apply a function over a list or vector.<br/>
Works like lapply(), but it tries to simplify the output to the most elementary data structure that is possible.

>- The tapply() Function
<br/> Apply a function over a ragged array.

>- The mapply() Function
<br/> Apply a function to multiple list or vector arguments.

## Parallel on server
For loop vs Apply

```{r, eval=FALSE}
for(i in 1:length(or_zip_name)){
  # output value of zipcode
  zipcode <- as.character(or_zip_name[i]) 
  # limit shapefile to particular zipcode
  zip_shape <- or_zip_map[or_zip_map$ZCTA5CE10 %in% zipcode, ]
  # convert to polygon
  zip_poly <-SpatialPolygons(zip_shape@polygons)

@@

One core vs Multiple cores
